Python as a machine learning language: experiences from the \textit{scikit-learn} project

### 2013-06-08 paper.tex
Minor rewording and a few remarks

The \textit{scikit-learn} project~\citep{pedregosa2011} is an open source
machine learning library for the Python programming language. The ambition of
the project is to provide  effective and long-established machine learning tools
within a programming environment that is accessible to non-machine learning
experts, efficient, and reusable in various scientific areas. In itself, the
project is a library of machine learning tools  rather than  a novel domain-
specific programming language. Among other things, the library includes
classical learning algorithms, model selection tools and preprocessing
procedures. From its core, the library has been designed to tie in with the set
of numeric and scientific packages centered around the Numpy and
SciPy~\citep{vanderwalt2011, varoquaux2013scipy} libraries. As such, \textit
{scikit-learn} is intended to be used within Python scripts and requires some
Python programming knowledge. At the time, it offers no command line or
graphical user interface.

@@

In \textit{scikit-learn}, we chose a representation of data that is as close as
possible to the matrix representation: datasets are encoded as NumPy
multidimensional arrays for dense data and as SciPy sparse matrices for sparse
data. While those may seem like bare and minimalistic data representations
when compared to more object-oriented
representations\footnote{Weka~\citep{hall2009weka} is an example of a toolkit-specific
object-oriented data representation.}, they bring the prime advantage
of allowing us to rely on all NumPy and SciPy vector operations -- often orders
of magnitude faster that the corresponding Python loops -- while allowing us at
the same time to keep the code short and readable. From a practical point of
view, these formats also provide a bunch of data loading and conversion tools
which make them very easy to use in many contexts\footnote{For tasks where the
inputs are text files or semi-structured objects, we provide \textit{vectorizer}
objects that efficiently convert such data to the NumPy or SciPy formats.}.
Finally, this design choice has also been motivated by the fact that, given
their pervasiveness in many other scientific Python packages, many scientific
users of Python are likely to be already familiar with those formats, hence helping
them to familiarize with the project.

@@

As an example, let us consider a supervised learning task using logistic regression.
Given the API defined above, solving this problem is as simple as:
\begin{center}
\texttt{clf = LogisticRegression(penalty="l1").fit(X\_train, y\_train)}
\end{center}
In this snippet, a \texttt{LogisticRegression} estimator is first initialized by
setting the \texttt{penalty} hyper-parameter to \texttt{l1}, hence enabling $L1$
regularization, and using default values for all other hyper-parameters. Upon
\texttt{fit}, the model is then learned from the training arrays \texttt{X\_train} and
\texttt{y\_train}. Also, since all estimators share the same interface, using a
different learning algorithm is as simple as replacing the constructor. For
example, to build a Random Forest on the same training data, one would simply
have to replace \texttt{LogisticRegression(penalty="l1")} in the snippet above by
\texttt{RandomForestClassifier()}. Let us also note that the snippet relies on the
fact that the \texttt{fit} method always returns the object it was called
on\footnote{I.e., the \texttt{self} argument it was called on.}, hence making
method chaining  possible.

>>>>

The \textit{scikit-learn} project~\citep{pedregosa2011} is an open source
machine learning library for the Python programming language. The ambition of
the project is to provide  effective and long-established machine learning tools
within a programming environment that is accessible to non-machine learning
experts, efficient, and reusable in various scientific areas. In itself, the
project is a library of machine learning tools  rather than  a novel domain-
specific programming language. Among other things, the library includes
classical learning algorithms, model selection tools and preprocessing
procedures. From its core, the library has been designed to tie in with the set
of numeric and scientific packages centered around the Numpy and
SciPy~\citep{vanderwalt2011, varoquaux2013scipy} libraries. As such, \textit
{scikit-learn} is intended to be used within Python scripts and requires some
Python programming knowledge. At the time, it offers no command line or
graphical user interface.
% XXX Gael: maybe we should say that it augments a general purpose
% language with machine learning concepts, like numpy augments it with
% array-computing concepts.

@@

In \textit{scikit-learn}, we chose a representation of data that is as close as
possible to the matrix representation: datasets are encoded as NumPy
multidimensional arrays for dense data and as SciPy sparse matrices for sparse
data. While those may seem rather unsophisticated data representations
when compared to more object-oriented
constructs, such as those used by Weka~\citep{hall2009weka}. However,
they bring the prime advantage
of allowing us to rely on all NumPy and SciPy vector operations -- often orders
of magnitude faster that the corresponding Python loops -- while allowing us at
the same time to keep the code short and readable. From a practical point of
view, these formats also provide a bunch of data loading and conversion tools
which make them very easy to use in many contexts\footnote{For tasks where the
inputs are text files or semi-structured objects, we provide \textit{vectorizer}
objects that efficiently convert such data to the NumPy or SciPy formats.}.
Finally, this design choice has also been motivated by the fact that, given
their pervasiveness in many other scientific Python packages, many scientific
users of Python are likely to be already familiar with those formats, hence helping
them to familiarize with the project.

@@

As an example, let us consider a supervised learning task using logistic regression.
Given the API defined above, solving this problem is as simple as:
% XXX: Gael: I think that we should break the following lines in 2: the
% 'return self' hidden in the call to fit can be confusing to the casual
% reader
\begin{center}
\texttt{clf = LogisticRegression(penalty="l1").fit(X\_train, y\_train)}
\end{center}
In this snippet, a \texttt{LogisticRegression} estimator is first initialized by
setting the \texttt{penalty} hyper-parameter to \texttt{l1}, hence enabling $L1$
regularization, and using default values for all other hyper-parameters. Upon
\texttt{fit}, the model is then learned from the training arrays \texttt{X\_train} and
\texttt{y\_train}. Also, since all estimators share the same interface, using a
different learning algorithm is as simple as replacing the constructor. For
example, to build a Random Forest on the same training data, one would simply
have to replace \texttt{LogisticRegression(penalty="l1")} in the snippet above by
\texttt{RandomForestClassifier()}. Let us also note that the snippet relies on the
fact that the \texttt{fit} method always returns the object it was called
on\footnote{I.e., the \texttt{self} argument it was called on.}, hence making
method chaining  possible.
